The development of children's ability to track and predict turn structure in conversation

### 2016-06-27 paper/CasillasFrank-Anticip.tex
separated supplementary materials per editor's request

The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot treat individual languages as factors. Gaze behavior might be best for non-native languages that have the most structural overlap with participants' native language: English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \citep{carlson2005}. We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age (3, 4, 5; numeric; intercept as age=0), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with two-way interactions between gap duration and the other simple fixed effects (language condition, transition type, and age) and a three-way interaction between language condition, transition type, and age.  We included the two-way interactions with gap duration in case the effect of inter-turn silence changes with age or linguistic cueing (e.g., if children older children rely less on silence as a cue)\footnote{We test these two-way interactions with gap duration in all of the models reported in this paper. Higher-order interactions with gap duration usually resulted in model non-convergence due to distributional sparsity when three or more predictor values were considered.}. We also included random effects of item (turn transition) and participant, with maximal random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report in this paper are all qualitatively unchanged by the exclusion of their random slopes. We have left the random slopes in because of minor participant-level variation in the predictors modeled.}

@@

Our primary analysis (above) makes the assumption that participants' eye movements generally follow the turn structure of the stimulus, i.e., that participants track the current speaker and switch their gaze to the upcoming speaker near turn transitions. As just described, based on this assumption, we used linear mixed effects regressions to see how anticipatory looking is affected by aspects of participant group (age) and stimulus (e.g., transition type, language condition). But what if the assumption that participants generally track turn structure were wrong? Could these results have emerged if participants' eye movements were \textit{not} linked to turn structure? For example, if participants were randomly looking back and forth between the two speakers, we might still find some anticipatory switching by chance. To test whether our primary results (the regression output above) could have arisen from random switching, not linked to turn structure, we conducted a secondary analysis comparing participants' anticipatory gaze at real and randomly shuffled points of turn transition.

We conducted this analysis by running the same regression models on participants' eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows. This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (see Figure \ref{fig:shuffling}; ``analysis window'' as shown in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described above) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data (our random baseline dataset) with the same statistical models we used for the original dataset (Table \ref{tab:E1-models}). Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each permutation.

@@

The random shuffling procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movement data. We created 5,000 permutations of the original turn transitions (step 1 above), thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data (step 2 above). Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other within a single iteration), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped).\footnote{This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech, which maximized the possibility of finding a difference from the baseline measure.} Pooled together, the anticipatory gaze datasets yielded an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline.

Using this technique we compared participants' anticipatory switches at turn transition windows to their anticipatory switches over the stimulus as a whole. If participants looked randomly back and forth between the speakers, we would have seen similar patterns in both cases. Rather than simply comparing participants' overall anticipatory switch rates with real and random transition windows, we estimated the likelihood that each of the predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) could have arisen with random gaze switching: we ran identical statistical models on the real and randomly permuted data sets. This tells us not only whether participants' switches were above chance, but whether the specific underlying effects of their anticipatory gaze patterns (e.g., the effect of language condition) were above that expected by chance. Because these analyses are complex and secondary to the main results, we report their full details in \ref{sec:permutation}.

Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking. For the children's data, the original \textit{z}-values for language condition, transition type, the age-language condition interaction, the transition type-gap duration interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values from models of the randomly permuted data (99.3\%, 99.1\%, 98.9\%, 97\%, and 96\%, respectively, all \textit{p}$<$.05). Similarly, the adults' data showed significant differentiation from the randomly permuted data for all three significant predictors from the real transition dataset. Transition type, the interaction between transition type and gap duration, and the interaction between language condition and transition type showed \textit{z}-values that exceeded 100\%, 99.8\%, and 95\% of random \textit{z}-values, respectively (all \textit{p}$\leq$.05). See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking. For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than active anticipation of an upcoming response. We address this hypothesis about ``turn-transition'' gaze switches vs. ``boredom'' gaze switches in \ref{sec:boredlooks}}

@@

The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random gaze switching (Figure \ref{fig:E1-randvsreal}). To further explore this effect, we compared the effect of language condition across age groups: using the permuted datasets described above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data. Then, within each permutation, we made pairwise comparisons of the average difference scores across participant age groups. This process yielded a distribution of random permutation-based difference scores that we could then compare to the difference score in the actual data. Details are given in \ref{sec:pairwisedev}.

These analyses revealed that, while 3- and 4-year olds showed similarly-sized effects of language condition, 5-year-olds had a significantly smaller effect of language condition, compared to both younger age groups. The difference in the language condition effect between 5-year-olds and 3-year-olds was greater than would be expected by chance (99.52\% of the randomly permuted data sets; \textit{p}$<$.01). Similarly, the difference in the language condition effect between 5-year-olds and 4-year-olds was greater than would be expected by chance (99.96\% of the data sets; \textit{p}$<$.001). See Figure \ref{fig:E1-lgageinteraction} for each difference score distribution.

@@

Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions. Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (\ref{sec:permutation}; Figure \ref{fig:E1-ChiTs}). These data show a number of important features that bear on our questions of interest.

First, both adults' and children's anticipations were strongly affected by transition type. Both groups made more anticipatory switches after hearing questions, compared to non-questions, especially for the English stimuli compared to the non-English stimuli. Overall, participants made few anticipatory switches after non-questions, even in the English videos when they had full linguistic access. Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{torreira2015, magyari2012, de-ruiter2006}. The current results add a new dimension to our understanding of how listeners make predictions about turn ends: both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to imminent responses.

@@

Finally, children and adults made more anticipatory switches in transitions with longer inter-turn gaps, though this effect was limited to non-question turns (Table \ref{tab:E1-models}). This finding suggests that gap duration indeed serves as a cue to upcoming turn structure; while short gaps may be perceived as within-turn pauses \citep{mannel2009}, long gaps could instead be indicative of between-turn pauses (where speaker transition occurs). Participants might use long silences to retroactively assign turn boundaries and anticipate speaker switches that were otherwise not anticipated (in this case, because the preceding turn was not a question). An alternative explanation for gap duration effects is that longer inter-turn gaps result in longer analysis windows, which yields more time for participants to make an anticipatory gaze. However, if participants are generally more likely to make a switch at question transitions, as our results suggest, we would expect that longer gaps would benefit questions more than non-questions---the opposite pattern from what the data show here. We take this as evidence that inter-turn silence may be most useful when participants have limited ability to make predictions about upcoming speaker transitions.

@@

We used three pairs of puppets for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: `robots', `merpeople', and `party-goers' (Figure 8). We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the partial cue conditions (`riding bikes', `rainy days', and `the library'). We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total; 6 videos per experiment version). We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.

@@

There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec. The remaining turn transitions had more questions than non-questions (N$=$47 and N$=$38, respectively), with transitions distributed somewhat evenly across conditions (keeping in mind that there were three \textit{normal} videos and only one video for each partial cue condition in each experiment version): \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19). Inter-turn gaps for questions (mean$=$366, median$=$438, stdev$=$138 msec) were longer than those for non-questions (mean$=$305, median$=$325, stdev$=$94 msec) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321, stdev$=$130 msec), \textit{words only} (mean$=$347, median$=$369, stdev$=$ 115 msec), \textit{prosody only} (mean$=$365, median$=$369, stdev$=$104 msec), and \textit{no words} (mean$=$319, median$=$329, stdev$=$136 msec).

@@

Using the same technique described in Experiment 1, we created and modeled random permutations of participants' anticipatory gaze switches. These analyses revealed that the significant predictors from models of the original, turn-related data were unlikely to be explained by random looking. In the children's data, the original model's \textit{z}-values for gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 93\% of the randomly permuted \textit{z}-values (95.6\%, 94\%, and 93.3\%, respectively, \textit{p}$=$.04, .06, and .07). Similarly, the adults' data showed significant differentiation from the randomly permuted data for the effect of language condition (\textit{words only} speech; greater than 98.3\% of random \textit{z}-values, \textit{p}$<$.02). See Section \ref{sec:permutation} for more information on each predictor's random permutation distribution.

@@

We also tested the two baseline linguistic conditions against each other---\textit{no speech} and \textit{normal speech}---to find out when linguistic information made a difference in children's anticipations. Because, as we have seen, children primarily show linguistic effects in question-answer turn transitions, we investigated the use of linguistic cues across age by testing anticipation separately for question and non-question turns. Compared to the \textit{no speech} condition, children made significantly more anticipatory switches in the \textit{normal} speech condition for questions at ages 6, 4, and 3, and also marginally at age 2 (6-year-olds: \textit{t}(36.919)$=$3.8019, \textit{p}$<$.001; 4-year-olds: \textit{t}(41.449)$=$2.9777, \textit{p}$<$.01; 3-year-olds: \textit{t}(35.724)$=$2.4286, \textit{p}$<$.05; 2-year-olds: \textit{t}(41.078)$=$1.8018, \textit{p}$=$.079). Children's anticipatory switches for questions did not differ in the \textit{no speech} and \textit{normal} speech conditions at ages 5 or 1 (5-year-olds: \textit{t}(29.406)$=$1.2783, \textit{p}$=$.211; 1-year-olds: \textit{t}(35.907)$=$0.4961, \textit{p}$=$.623). In contrast, children's anticipatory switch rates for non-question turns were not significantly different between the \textit{no speech} and \textit{normal} speech conditions at any age (all \textit{p}$>$0.09). Consistent with the regression results, children were more likely to show an effect of linguistic content as they got older, but only for question transitions.

The regression models for the children's data also revealed two significant interactions with age. The first was a significant interaction of age and language condition (for \textit{prosody only} compared to the \textit{no speech} reference level), suggesting a different age effect between the two linguistic conditions. As in Experiment 1, we explored each age interaction by extracting an average difference score over participants for the effect of language condition (\textit{no speech} vs. \textit{prosody only}) within each random permutation of the data, making pairwise comparisons between the six age groups. These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six compared to the \textit{no speech} baseline (with difference scores greater than 95\% of the random data scores; \textit{p}$<$.05). See Figure \ref{fig:E2-lgageinteraction} for these \textit{prosody only} difference score distributions.

The second age-based interaction was a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} baseline). We again created pairwise comparisons of the average difference scores for the transition type-language condition interaction across age groups in each random permutation of the data, finding that the effect of transition type in the \textit{normal} speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p}$<$.05). See Figure \ref{fig:E2-lgagetypeinteraction} for these \textit{normal} speech difference score distributions.

@@

As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations. Participants made anticipatory gaze switches at above-chance rates starting at age two for both questions and non-questions. Longer gaps had a broader impact on participants' anticipations in this experiment; we saw that, overall, longer inter-turn gaps resulted in more anticipatory switches, with the \textit{no speech} condition showing equal or stronger effects of gap duration than all other conditions.

As before, participants made far more anticipations for questions than for non-question turns---at least for those two years old and older. But these effects were different for the conditions with partial linguistic information: \textit{prosody only} and \textit{words only}. In the \textit{prosody only} condition, performance was initially low for young children and increased significantly with age. In the \textit{words only} condition, children age two and older showed robust switching for questions (much like in \textit{normal} speech), but never rose above chance for non-question turns (Figure \ref{fig:E2-randvsreal}), with no significant differences from the \textit{no speech} baseline. These findings do not support an early role for prosody or lexical information alone in children's spontaneous predictions about turn structure. They also give no support for the idea that lexical information is sufficient on its own to support children's anticipatory switching. They do underscore the developing relationship between the use of linguistic cues, inter-turn silence, and speech act (transition type) in spontaneous predictions about upcoming turn structure.

@@

Children begin to develop conversational turn-taking skills long before their first words emerge \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}. As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure. Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors. In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that, when spontaneous predictions begin, they are primarily driven by participants' expectation of an immediate response in the next turn (e.g., after questions). In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither signal alone was sufficient to support increased anticipatory gaze. We also found no early advantage for prosody over lexicosyntax; children's anticipatory switch rates in the \textit{prosody only} condition were initially low, but showed significant gains by age five. We discuss these findings with respect to the role of linguistic cues and inter-turn silence for predicting upcoming turn structure, the importance of questions in predictions about conversation, and children's developing competence as conversationalists.

@@

Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody is still under debate \citep{duncan1972, ford1996, torreira2015}. Knowing that children comprehend more about prosody than lexicosyntax early on (see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation. Our results suggest that, on the contrary, exclusively presenting prosodic information to children limits their spontaneous predictions about upcoming turn structure until age five.

Thus, using prosody alone to accurately predict turn boundaries in conversation appears to be difficult for adults and children. Prosodic information is continuous, multidimensional, and indexically complex---it encodes syntactic structure, speech act, and extralinguistic information, sometimes simultaneously, without clear one-to-one mappings between form and meaning \citep{cutler1997, shriberg1998, lammertink2015}. For these reasons, prosodic information alone may not be enough to both (a) make precise temporal predictions about turn structure, and (b) identify question turns, which otherwise appear to drive anticipatory gaze switching. Therefore, although children show early facility with prosodic discrimination \citep{nazzi2003, soderstrom2003, johnson2001, jusczyk1995, morgan1995, mehler1988}, using prosodic knowledge for turn prediction may generally be too difficult without additional information from lexical or syntactic cues.

>>>>

The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot treat individual languages as factors. Gaze behavior might be best for non-native languages that have the most structural overlap with participants' native language: English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \citep{carlson2005}. We would need multiple items from each of the languages to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age (3, 4, 5; numeric; intercept as age=0), and duration of the inter-turn gap (seconds, e.g., 0.441) as predictors, with two-way interactions between gap duration and the other simple fixed effects (language condition, transition type, and age) and a three-way interaction between language condition, transition type, and age.  We included the two-way interactions with gap duration in case the effect of inter-turn silence changes with age or linguistic cueing (e.g., if children older children rely less on silence as a cue).\footnote{We test these two-way interactions with gap duration in all of the models reported in this paper. Higher-order interactions with gap duration usually resulted in model non-convergence due to distributional sparsity when three or more predictor values were considered, so we did not include them.} We also included random effects of item (turn transition) and participant, with maximal random slopes of condition, transition type, and their interaction for participants \citep{barr2013}.\footnote{The models we report in this paper are all qualitatively unchanged by the exclusion of their random slopes. We have left the random slopes in because of minor participant-level variation in the predictors modeled.}

@@

Our primary analysis (above) makes the assumption that participants' eye movements generally follow the turn structure of the stimulus, i.e., that participants track the current speaker and switch their gaze to the upcoming speaker near turn transitions. As just described, based on this assumption, we used linear mixed effects regressions to see how anticipatory looking is affected by aspects of participant group (e.g., age) and stimulus (e.g., transition type, language condition). But what if the assumption that participants generally track turn structure were wrong? Could these results have emerged if participants' eye movements were \textit{not} linked to turn structure? For example, if participants were randomly looking back and forth between the two speakers, we might still find some anticipatory switching by chance. To test whether our primary results (the regression output above) could have arisen from random switching we conducted a secondary analysis comparing participants' anticipatory gaze at real and randomly shuffled points of turn transition.

We conducted this analysis by running the same regression models on participants' eye-tracking data, only this time calculating their anticipatory gaze switches with respect to randomly permuted turn transition windows. This process involved: (1) randomizing the order and temporal placement of the analysis windows within each stimulus (Figure \ref{fig:shuffling}; ``analysis window'' is as shown in Figure \ref{fig:criterion}) to randomly redistribute the analysis windows across the eye-tracking signal, (2) re-running each participant's eye tracking data through switch identification (described above) on each of the randomly permuted analysis windows, and (3) modeling the anticipatory switches from the randomly permuted data (our random baseline dataset) with the same statistical models we used for the original dataset (Table \ref{tab:E1-models}). Importantly, although the onset time of each transition was shuffled within the eye-tracking signal, the other intrinsic properties of each turn transition (e.g., prior speaker identity, transition type, gap duration, language condition, etc.) stayed constant across each permutation.

@@

The random shuffling procedure de-links participants' gaze data from the turn structure in the original stimulus, thereby allowing us to compare turn-related (original) and non-turn-related (randomly permuted) looking behavior using the same eye movement data. We created 5,000 permutations of the original turn transitions, thereby creating 5,000 anticipatory gaze datasets with randomly de-linked gaze data. Because the randomly shuffled turn transitions could occur anywhere in the stimulus (so long as they didn't overlap each other within a single iteration), the resulting turn-transition windows collectively covered the entire stimulus---during speech and silence, during speaker change and speaker continuation, and during all turn transitions in the stimulus, even those excluded in the original analyses (e.g., because they were overlapped). This technique crucially differs from that used by Keitel and colleagues (2013, 2015), which tests anticipatory gaze at turn transitions against anticipatory gaze during speech. Pooled together, our 5,000 anticipatory gaze datasets yielded an average anticipatory switch rate for each participant over all possible starting points in the stimuli: a random baseline. Using this technique we compared participants' anticipatory switches at turn transition windows to their anticipatory switches over the stimulus as a whole. If participants looked randomly back and forth between the speakers, we would have seen similar patterns in both cases.

Rather than simply comparing participants' overall anticipatory switch rates with real and random transition windows, we estimated the likelihood that each of the predictor effects in the original data (e.g., the effect of language condition; Table \ref{tab:E1-models}) could have arisen with random gaze switching: we ran identical statistical models on the real and randomly permuted data sets. This tells us not only whether participants' switches were above chance, but whether the specific underlying effects of their anticipatory gaze patterns (e.g., the effect of language condition) were above that expected by chance. Because these analyses are complex and secondary to the main results, we report their full details in Supplementary material A.

Our baseline analyses revealed that none of the significant predictors from models of the original, turn-related data can be explained by random looking. For the children's data, the original \textit{z}-values for language condition, transition type, the age-language condition interaction, the transition type-gap duration interaction, and the language condition-transition type interaction were all greater than 95\% of \textit{z}-values from models of the randomly permuted data (99.3\%, 99.1\%, 98.9\%, 97\%, and 96\%, respectively, all \textit{p}$<$.05). Similarly, the adults' data showed significant differentiation from the randomly permuted data for all three significant predictors from the real transition dataset. Transition type, the interaction between transition type and gap duration, and the interaction between language condition and transition type showed \textit{z}-values that exceeded 100\%, 99.8\%, and 95\% of random \textit{z}-values, respectively (all \textit{p}$\leq$.05). See Supplementary material A for more information on each predictor's random permutation distribution.\footnote{This baseline analysis tests ``random looking'' against ''turn-driven looking'', but it does not test subtypes of turn-driven looking. For example, children might switch their gaze from the current speaker to the addressee out of boredom with the ongoing speech rather than from active anticipation of an upcoming response. We address this hypothesis about ``boredom'' gaze switches vs. ``turn-transition'' gaze switches in Supplementary material C.}

@@

The models reported above revealed a significant interaction of age and language condition (Table \ref{tab:E1-models}) that was unlikely be due to random gaze switching (Figure \ref{fig:E1-randvsreal}). To further explore this effect, we compared the effect of language condition across age groups: using the permuted datasets described above, we extracted the average difference score for the two language conditions (English minus non-English) for each participant, computing an overall average for each random permutation of the data. Then, within each permutation, we made pairwise comparisons of the average difference scores across participant age groups. This process yielded a distribution of random permutation-based difference scores that we could then compare to the difference score in the actual data. Details are given in Supplementary material B.

These analyses revealed that, while 3- and 4-year olds showed similarly-sized effects of language condition, 5-year-olds had a significantly smaller effect of language condition, compared to both younger age groups. The difference in the language condition effect between 5-year-olds and 3-year-olds was greater than would be expected by chance (99.52\% of the randomly permuted data sets; \textit{p}$<$.01). Similarly, the difference in the language condition effect between 5-year-olds and 4-year-olds was greater than would be expected by chance (99.96\% of the data sets; \textit{p}$<$.001). See Supplementary material Figure B.1 for each difference score distribution.

@@

Children and adults spontaneously tracked the turn structure of the conversations, making anticipatory gaze switches at an above-chance rate across all ages and conditions. Children's anticipatory gaze rates were affected by language condition, transition type, age, and gap duration (Table \ref{tab:E1-models}), none of which could be explained by a baseline of random gaze switching (see Supplementary material Figure A.1). These data show a number of important features that bear on our questions of interest.

First, both adults' and children's anticipations were strongly affected by transition type. Both groups made more anticipatory switches after hearing questions, compared to non-questions, especially for the English stimuli compared to the non-English stimuli. Overall, participants made few anticipatory switches after non-questions, even in the English videos when they had full linguistic access. Prior work using online, metalinguistic tasks has shown that participants can use linguistic cues to accurately predict upcoming turn ends \citep{bogelstorreira2015, magyari2012, de-ruiter2006}. The current results add a new dimension to our understanding of how listeners make predictions about turn ends: both children and adults spontaneously monitor the linguistic structure of unfolding turns for cues to imminent responses.

@@

Finally, children and adults made more anticipatory switches in transitions with longer inter-turn gaps, though this effect was limited to non-question turns (Table \ref{tab:E1-models}). This finding suggests that gap duration indeed serves as a cue to upcoming turn structure; while short gaps may be perceived as within-turn pauses \citep{mannel2009}, long gaps could instead be indicative of between-turn pauses (where speaker transition occurs). Participants might use long silences to retroactively assign turn boundaries and anticipate speaker switches that were otherwise not anticipated (in this case, because the preceding turn was not a question). An alternative explanation for effects of gap duration is that longer inter-turn gaps result in longer analysis windows, which gives participants more time to make an anticipatory gaze. However, if participants are generally more likely to make a switch at question transitions (as our results suggest), and if question-driven switches aren't already at ceiling when gaps are short, we would expect that longer gaps would benefit questions more than non-questions---the opposite pattern from what the data show here. We take this as evidence that inter-turn silence may be most useful when participants have limited ability to make predictions about upcoming speaker transitions.

@@

We used three pairs of puppets for the \textit{normal} condition---`red', `blue' and `yellow'---and one pair of puppets for each partial cue condition: `robots', `merpeople', and `party-goers' (Figure \ref{fig:puppets}). We randomly assigned half of the conversation topics (`birthday cake', `pets', and `breakfast') to the \textit{normal} condition, and half to the partial cue conditions (`riding bikes', `rainy days', and `the library'). We then created three versions of the experiment, so that each of the six puppet pairs was associated with three different conversation topics across the different versions of the experiment (18 videos in total; 6 videos per experiment version). We ensured that the position of the talkers (left and right) was counterbalanced in each version by flipping the video and audio channels as needed.

@@

There were 85 turn transitions for analysis after excluding transitions longer than 550 msec and shorter than 90 msec. The remaining turn transitions had more questions than non-questions (N$=$47 and N$=$38, respectively), with transitions distributed somewhat evenly across conditions, keeping in mind that there were three \textit{normal} videos and only one video for each partial cue condition in each experiment version: \textit{normal} (N$=$36), \textit{words only} (N$=$13), \textit{prosody only} (N$=$17), and \textit{no speech} (N$=$19). Inter-turn gaps for questions (mean$=$366, median$=$438, stdev$=$138 msec) were longer than those for non-questions (mean$=$305, median$=$325, stdev$=$94 msec) on average, but gap duration was overall comparable across conditions: \textit{normal} (mean$=$334, median$=$321, stdev$=$130 msec), \textit{words only} (mean$=$347, median$=$369, stdev$=$ 115 msec), \textit{prosody only} (mean$=$365, median$=$369, stdev$=$104 msec), and \textit{no words} (mean$=$319, median$=$329, stdev$=$136 msec).

@@

Using the same technique described in Experiment 1, we created and modeled random permutations of participants' anticipatory gaze switches. These analyses revealed that the significant predictors from models of the original, turn-related data were unlikely to be explained by random looking. In the children's data, the original model's \textit{z}-values for gap duration, the two-way interaction of age and language condition (\textit{prosody only}) and the three-way interaction of age, transition type, and language condition (\textit{normal} speech) were all greater than 93\% of the randomly permuted \textit{z}-values (95.6\%, 94\%, and 93.3\%, respectively, \textit{p}$=$.04, .06, and .07). Similarly, the adults' data showed significant differentiation from the randomly permuted data for the effect of language condition (\textit{words only} speech; greater than 98.3\% of random \textit{z}-values, \textit{p}$<$.02). See Supplementary material A for more information on each predictor's random permutation distribution.

@@

We also tested the two baseline linguistic conditions against each other---\textit{no speech} and \textit{normal speech}---to find out when linguistic information made a difference in children's anticipations. Because, as we have seen, children primarily show linguistic effects in question-answer turn transitions, we investigated the use of linguistic cues across age by testing anticipation separately for question and non-question turns. Compared to the \textit{no speech} condition, children made significantly more anticipatory switches in the \textit{normal} speech condition for questions at ages 6, 4, and 3, and also marginally at age 2 (6-year-olds: \textit{t}(36.919)$=$3.8019, \textit{p}$<$.001; 4-year-olds: \textit{t}(41.449)$=$2.9777, \textit{p}$<$.01; 3-year-olds: \textit{t}(35.724)$=$2.4286, \textit{p}$<$.05; 2-year-olds: \textit{t}(41.078)$=$1.8018, \textit{p}$=$.079). Children's anticipatory switches for questions did not significantly differ in the \textit{no speech} and \textit{normal} speech conditions at ages 5 or 1 (5-year-olds: \textit{t}(29.406)$=$1.2783, \textit{p}$=$.211; 1-year-olds: \textit{t}(35.907)$=$0.4961, \textit{p}$=$.623). In contrast, children's anticipatory switch rates for non-question turns were not significantly different between the \textit{no speech} and \textit{normal} speech conditions at any age (all \textit{p}$>$0.09). Thus, consistent with the regression results, children were more likely to show an effect of linguistic content as they got older, but only for question transitions.

The regression models for the children's data also revealed two significant interactions with age. The first was a significant interaction of age and language condition (for \textit{prosody only} compared to the \textit{no speech} reference level), suggesting a different age effect between the two linguistic conditions. As in Experiment 1, we explored each age interaction by extracting an average difference score over participants for the effect of language condition (\textit{no speech} vs. \textit{prosody only}) within each random permutation of the data, making pairwise comparisons between the six age groups. These tests revealed that children's anticipation in the \textit{prosody only} condition significantly improved at ages five and six compared to the \textit{no speech} baseline (with difference scores greater than 95\% of the random data scores; \textit{p}$<$.05). See Supplementary material Figure B.2 for these \textit{prosody only} difference score distributions.

The second age-based interaction was a three-way interaction of age, transition type, and language condition (for \textit{normal} speech compared to the \textit{no speech} baseline). We again created pairwise comparisons of the average difference scores for the transition type-language condition interaction across age groups in each random permutation of the data, finding that the effect of transition type in the \textit{normal} speech condition became larger with age, with significant improvements by age 4 over ages 1 and 2 (99.9\% and 98.86\%, respectively), by age 5 over age 4 (97.54\%), and by age 6 over ages 1, 2, and 5 (99.5\%, 97.36\%, and 95.04\%), all significantly different from chance (\textit{p}$<$.05). See Supplementary material Figure B.3 for these \textit{normal} speech difference score distributions.

@@

As in Experiment 1, children and adults spontaneously tracked the turn structure of the conversations. Participants made anticipatory gaze switches at above-chance rates starting at age two for both questions and non-questions. Longer gaps had a broader impact on participants' anticipations in this second experiment; we saw that, overall, longer inter-turn gaps resulted in more anticipatory switches, with the \textit{no speech} condition showing equal or stronger effects of gap duration than all other conditions.

As before, participants made far more anticipations for questions than for non-question turns---at least for those two years old and older. But these effects were different for the conditions with partial linguistic information: \textit{prosody only} and \textit{words only}. In the \textit{prosody only} condition, performance was initially low for young children and increased significantly with age. In the \textit{words only} condition, children age two and older showed robust switching for questions (much like in \textit{normal} speech), but never rose above chance for non-question turns (Figure \ref{fig:E2-randvsreal}), with no significant differences from the \textit{no speech} baseline. These findings do not support an early role for prosody or lexical information alone in children's spontaneous predictions about turn structure. They also give no support for the idea that lexical information is sufficient on its own to support children's anticipatory switching. They do underscore the developing relationship between the online use of linguistic cues, inter-turn silence, and speech act in spontaneous predictions about upcoming turn structure.

@@

Children begin to develop conversational turn-taking skills long before their first words emerge \citep{bateson1975, hilbrink2015, jaffe2001, snow1977}. As they become fast and knowledgeable language users, they also become able to make accurate predictions about upcoming turn structure. Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors. In two experiments investigating children's anticipatory gaze to upcoming speakers, we found evidence that turn prediction develops early in childhood and that, when spontaneous predictions begin, they are primarily driven by participants' expectation of an immediate response in the next turn (e.g., after questions). In making predictions about upcoming turn structure, children used a combination of lexical and prosodic cues; neither signal alone was sufficient to support increased anticipatory gaze. We also found no early advantage for prosody over lexicosyntax; children's anticipatory switch rates in the \textit{prosody only} condition were initially low, but showed significant gains by age five. We discuss these findings with respect to the role of linguistic processing and inter-turn silence for predicting upcoming turn structure, the importance of questions in predictions about conversation, and children's developing competence as conversationalists.

@@

Prior work with adults has found a consistent role for lexicosyntax in predicting upcoming turn structure \citep{de-ruiter2006, magyari2012}, whereas the role of prosody is still under debate \citep{duncan1972, ford1996, bogelstorreira2015}. Knowing that children comprehend more about prosody than lexicosyntax early on (see \citealp{speer2009} for a review), we thought it possible that young children would instead show an advantage for prosody in their predictions about turn structure in conversation. Our results suggest that, on the contrary, exclusively presenting prosodic information to children limits their spontaneous predictions about upcoming turn structure until age five.

Thus, using prosody alone to accurately predict turn boundaries in conversation appears to be difficult for adults and children. Prosodic information is continuous, multidimensional, and can index multiple meanings at once---it encodes syntactic structure, speech act, and extralinguistic information without clear one-to-one mappings between form and meaning \citep{cutler1997, shriberg1998, lammertink2015}. For these reasons, prosodic information alone may not be enough for young children to easily make precise temporal predictions about turn structure, and identify question turns in unfolding speech. Therefore, although children show early facility with prosodic discrimination \citep{nazzi2003, soderstrom2003, johnson2001, jusczyk1995, morgan1995, mehler1988}, using prosodic knowledge for turn prediction may be difficult without additional information from lexical or syntactic cues.
